### 2021-05-28 thesis/tex/DCtheory.tex
update

\documentclass[draft=True]{thesis}
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx, slashed, siunitx}
\usepackage[utf8]{inputenc}
\usepackage{xr-hyper}
\input{commands}
\begin{document}
\section{DeepCore}
In this part, we use the publically available DeepCore data sample~\cite{DC2019data} which is an updated version of what was used by the 
IceCube collaboration in a $\nu_\mu$ disapprearance analysis~\cite{DC2018mudisappearance}.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.
\section{PINGU}
The methodology behind the PINGU simulations are the same as with our DeepCore study~\ref{ch:DCmethod}. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modelled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU MC is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We generate "data" by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

\begin{align}\label{eq:PINGUparams}
    &\Delta m^2_{21} =  \SI{7.42e-5}{\electronvolt^2},\hspace{0.5em} \dm =  \SI{2.517e-3}{\electronvolt^2}, \nonumber \\
    &\theta_{12} = \SI{33.44}{\degree},\hspace{1em} \theta_{13} = \SI{8.57}{\degree},\hspace{1em} \theta_{23} = \SI{49.2}{\degree}, \hspace{1em} \delta_\text{CP} = 0\,.
\end{align}
\bibliographystyle{nature}
\bibliography{ref.bib}
\end{document}

>>>>

% \documentclass[draft=True]{thesis}
% \usepackage[margin=0.5in]{geometry}
% \usepackage{graphicx, slashed, siunitx}
% \usepackage[utf8]{inputenc}
% \usepackage{xr-hyper}
% \input{commands}
% \begin{document}
\section{DeepCore}\label{ch:DCmethod}
In this part, we use the publically available DeepCore data sample~\cite{DC2019data} which is an updated version of what was used by the 
IceCube collaboration in a $\nu_\mu$ disapprearance analysis~\cite{DC2018mudisappearance}.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.
\section{PINGU}
The methodology behind the PINGU simulations are the same as with our DeepCore study~. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modelled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU Monte Carlo is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We generate "data" by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

\begin{align}\label{eq:PINGUparams}
    &\Delta m^2_{21} =  \SI{7.42e-5}{\electronvolt^2},\hspace{0.5em} \dm =  \SI{2.517e-3}{\electronvolt^2}, \nonumber \\
    &\theta_{12} = \SI{33.44}{\degree},\hspace{1em} \theta_{13} = \SI{8.57}{\degree},\hspace{1em} \theta_{23} = \SI{49.2}{\degree}, \hspace{1em} \delta_\text{CP} = 0\,.
\end{align}
% \bibliographystyle{nature}
% \bibliography{ref.bib}
% \end{document}

### 2021-05-28 thesis/tex/ICintro.tex
update

\documentclass{thesis}
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx, slashed, siunitx, subcaption}
\usepackage[utf8]{inputenc}
\usepackage{xr-hyper}
\input{commands}
\begin{document}
\section{Neutrino detection}
We always observe neutrinos indirectly through their associated charged lepton. Regardless of the type of interaction (charged current via the $W$ boson, or neutral current
via the $Z$), a charged lepton exits with altered properties. The lepton is then detected, and the properties of the neutrino involved in the 
interaction is then deduced. This deduction is obviously imperfect, and this introduces complexities that we will handle in Ch.~\ref{ch:ICmethod}.

@@

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than other.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single digit \si{\GeV}. Due to 
this part being situated deep within the ice, it is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}\label{fig:array}
    \includegraphics[width=0.5\textwidth]{figures/icecube2.png}\caption{View of the full IceCube array}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The "Precision IceCube Next Generation Upgrade" is an upgrade that will 
supplement DeepCore, i.e. boosing the capabilites of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
IceCube and DeepCore bounds. The PINGU simulations have the same structure as the DeepCore data, so any analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the DeepCore experiment.
\bibliographystyle{nature}
\bibliography{ref.bib}
\end{document}

>>>>

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \usepackage{graphicx, slashed, siunitx, subcaption}
% \usepackage[utf8]{inputenc}
% \usepackage{xr-hyper}
% \input{commands}
% \begin{document}
\section{Neutrino detection}
We always observe neutrinos indirectly through their associated charged lepton. Regardless of the type of interaction (charged current via the $W$ boson, or neutral current
via the $Z$), a charged lepton exits with altered properties. The lepton is then detected, and the properties of the neutrino involved in the 
interaction is then deduced. This deduction is obviously imperfect, and this introduces complexities that we will handle in Ch.~\ref{ch:ICmethod}.

@@

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than other.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single digit \si{\GeV}. Due to 
this part being situated deep within the ice, it is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}\label{fig:array}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosing the capabilites of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
IceCube and DeepCore bounds. The PINGU simulations have the same structure as the DeepCore data, so any analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the DeepCore experiment.

\begin{figure}[t]\label{fig:flux_aeff}
    \includegraphics[width=0.5\textwidth]{figures/flux.pdf}\hspace{-1cm}
    \includegraphics[width=0.5\textwidth]{figures/aeff.pdf}
    \vspace{-2.5cm}\caption{\emph{Left panel:} Interpolated South Pole atmospheric flux with data from~\cite{hondaArticle}.
    \emph{Right panel:} Interpolated IceCube effective area with data from~\cite{ICaeff}.}
\end{figure}

\subsection{Atmospheric neutrino flux}
Atmospheric neutrinos originates from cosmic rays composed of protons interacting with nuclei in the atmosphere.
These interactions ultimately produces pions, which decay as 
\begin{align}\label{eq:pion}
    &\pi^+ \to \mu^+ + \nm\,, \quad \pi^- \to \mu^- + \anm \nonumber \\
    &\pi^+ \to e^+ + \ne\,,\,\, \quad \pi^- \to e^- + \ane\,.
\end{align}
In the muonic decay channel, muons are emitted which will be detected by the IceCube detector. A part of the 
uncorrelated systematic error comes from this \emph{muon background}, i.e. events misclassified as muons from 
$\nm$ interactions rather than from pion decay. Moreover, the atmospheric flux is often associated with a large error.
In this work, we will use a flux normalization error of 24\%, and a zenith slope error of 4\%~\cite{hondapaper}.

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}
The flux data is binned in $\ztrue$. The fluxes are averaged over azimuthal direction and over solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\nu_{\bar{\tau}}$ are missing. Kaons can decay into neutral pions, which in turn 
can produce $\nt$, but this branching ratio is extremely small. Thus, we never have to use probabilities on the form 
$P_{\tau \beta}$, since we have no incoming atmospheric $\nu_\tau$ flux.

Interpolating the data yields makes us capable of returning all four necessary fluxes for a given true energy and true zenith.
The result is shown in Fig.~\ref{fig:flux_aeff}.

\begin{table}[h]\label{table:flux}
    \begin{center}
        \begin{tabular}{lcccccc}
            \hline \hline
            $\Etrue$ [\si{\GeV}] &$\phi_\mu$ &$\phi_{\bar{\mu}}$ &$\phi_e$ &$\phi_{\bar{e}}$ & $\ztrue$\\\\
            \hline
            27825 &  \SI{6.06e-12}{} &  \SI{3.17e-12}{} &  \SI{1.56e-13}{} &  \SI{1.04e-13}{} &   [-0.2, -0.1] \\
            247707 &  \SI{5.94e-16}{} &  \SI{2.92e-16}{} &  \SI{1.36e-17}{} &  \SI{8.12e-18}{} &   [-0.7, -0.6] \\
                22 &  \SI{3.33e-02}{} &  \SI{2.78e-02}{} &  \SI{9.57e-03}{} & \SI{7.15e-03}{} &   [-0.3, -0.2] \\
            432876 &  \SI{5.19e-17}{} &  \SI{2.32e-17}{} &  \SI{1.46e-18}{} & \SI{9.83e-19}{} &   [-1.1, -1.0] \\
            64280 &  \SI{1.58e-13}{} &  \SI{8.10e-14}{} &  \SI{3.49e-15}{} &  \SI{2.21e-15}{} &   [-0.4, -0.3] \\
            \hline \hline
        \end{tabular}
    \end{center}
    \caption{A selection of processed atmospheric South Pole fluxes from~\cite{hondaData} by Honda et al.~\cite{hondaArticle}.}
\end{table}

\subsection{Event reconstruction}
After an event has occurred, the IceCube algorithms process the data coming from the detector to \emph{reconstruct} the event. This means that, given the parameters recorded by the detector, what are their "true" values?
We are interested in two variables: the energy and the direction. Each event is tagged with a probable energy and zenith angle, called the recostructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
The collaboration then uses numerous sophisticated methods to backtrack the reconstructed parameters to the true parameters. So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true and reconstructed energy and zenith angle. The reconstructed parameters are what we are using to analyze the data (because this is what the detector actually sees), while the true parameters are used in the determination of that neutrino's "actual" flux and cross-section (because this is what nature sees).

How do we then translate between the reconstructed and true parameters? In this work, we are using two different methods, which are based on the form of data available to us. They will be outlined in Sec.\ref{ch:ICmethod} and Sec.~\ref{ch:DCmethod}.

% \bibliographystyle{nature}
% \bibliography{ref.bib}
% \end{document}

### 2021-05-28 thesis/tex/ICtheory.tex
update

\documentclass{thesis}
\usepackage[margin=0.5in]{geometry}
\input{commands}
\begin{document}
\begin{figure}[t]\label{fig:flux_aeff}
    \includegraphics[width=0.5\textwidth]{figures/flux.pdf}\hspace{-1cm}
    \includegraphics[width=0.5\textwidth]{figures/aeff.pdf}
    \vspace{-2.5cm}\caption{\emph{Left panel:} Interpolated South Pole atmospheric flux with data from~\cite{hondaArticle}.
    \emph{Right panel:} Interpolated IceCube effective area with data from~\cite{ICaeff}.}
\end{figure}

\section{Atmospheric neutrino flux}
Atmospheric neutrinos originates from cosmic rays composed of protons interacting with nuclei in the atmosphere.
These interactions ultimately produces pions, which decay as 
\begin{align}\label{eq:pion}
    \pi^+ &\to \mu^+ + \nm\,, \quad \pi^- \ to \mu^- + \anm \nonumber \\
    \pi^+ &\to e^+ + \ne\,, \quad \pi^- \ to e^- + \ane\,.
\end{align}
In the muonic decay channel, muons are emitted which will be detected by the IceCube detector. A part of the 
uncorrelated systematic error comes from this \emph{muon background}, i.e. events misclassified as muons from 
$\nm$ interactions rather than from pion decay. Moreover, the atmospheric flux is often associated with a large error.
In this work, we will use a flux normalization error of 24\%, and a zenith slope error of 4\%. In more sophisticated 
models, one can also include the pion to kaon flux uncertainty.

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}
The flux data is binned in $\ztrue$. The fluxes are averaged over azimuthal direction and over solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\nu_{\bar{\tau}}$ are missing. Kaons can decay into neutral pions, which in turn 
can produce $\nt$, but this branching ratio is extremely small. Thus, we never have to use probabilities on the form 
$P_{\tau \beta}$, since we have no incoming atmospheric $\nu_\tau$ flux.

Interpolating the data yields makes us capable of returning all four necessary fluxes for a given true energy and true zenith.
The result is shown in Fig.~\ref{fig:flux_aeff}.

\begin{table}[h]\label{table:flux}
    \begin{center}
        \begin{tabular}{lcccccc}
            \hline \hline
            $\Etrue$ [\si{\GeV}] &$\phi_\mu$ &$\phi_{\bar{\mu}}$ &$\phi_e$ &$\phi_{\bar{e}}$ & $\ztrue_{min}$ &$\ztrue_{max}$\\
            \hline
            27825 &  \SI{6.06e-12}{} &  \SI{3.17e-12}{} &  \SI{1.56e-13}{} &  \SI{1.04e-13}{} &   -0.2 &   -0.1 \\
            247707 &  \SI{5.94e-16}{} &  \SI{2.92e-16}{} &  \SI{1.36e-17}{} &  \SI{8.12e-18}{} &   -0.7 &   -0.6 \\
                22 &  \SI{3.33e-02}{} &  \SI{2.78e-02}{} &  \SI{9.57e-03}{} & \SI{7.15e-03}{} &   -0.3 &   -0.2 \\
            432876 &  \SI{5.19e-17}{} &  \SI{2.32e-17}{} &  \SI{1.46e-18}{} & \SI{9.83e-19}{} &   -1.1 &   -1.0 \\
            64280 &  \SI{1.58e-13}{} &  \SI{8.10e-14}{} &  \SI{3.49e-15}{} &  \SI{2.21e-15}{} &   -0.4 &   -0.3 \\
            \hline
        \end{tabular}
    \end{center}
    \caption{A selection of processed atmospheric South Pole fluxes from~\cite{hondaData} by Honda et al.~\cite{hondaArticle}.}
\end{table}

\section{Event reconstruction}
After an event has occurred, the IceCube algorithms process the data coming from the detector to \emph{reconstruct} the event. This means that, given the parameters recorded by the detector, what are their "true" values?
We are interested in two variables: the energy and the direction. Each event is tagged with a probable energy and zenith angle, called the recostructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
The collaboration then uses numerous sophisticated methods to backtrack the reconstructed parameters to the true parameters. So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true and reconstructed energy and zenith angle. The reconstructed parameters are what we are using to analyze the data (because this is what the detector actually sees), while the true parameters are used in the determination of that neutrino's "actual" flux and cross-section (because this is what nature sees).

How do we then translate between the reconstructed and true parameters? In this work, we are using two different methods, which are based on the form of data available to us.

\section{IceCube event count}\label{ch:ICmethod}
As the neutrinos have propagated the Earth, they arrive at the South Pole, where they interact in the ice to form their charged leptons. We now are interested in the effective area, i.e.~the detector area that the lepton "sees".
The effective area, $\Aeff$, depends on several parameters, some of them being detector physical volume, $\Etrue$,$\ztrue$ and the neutrino cross-section. Fortunately, the binned $\Aeff$ is provided to us by the collaboration~\cite{ICaeff}.
The data file has the following form

\begin{table}\label{table:aeff}
    \begin{tabular}{lrrrrr}
        \hline \hline
        $\Etrue_{min}$ [\si{\GeV}] &     $\Etrue_{max}$ [\si{\GeV}]&   $\ztrue_{min}$ &   $\ztrue_{max}$ &     $\Aeff$ [\si{\metre\squared}] \\
        \hline
             251 &      316 &  -0.92 &  -0.91 &   0.0174 \\
          794300 &  1000000 &  -0.80 &  -0.79 &  69.3600 \\
            3981 &     5012 &  -0.78 &  -0.77 &   3.1490 \\
            1585 &     1995 &  -0.07 &  -0.06 &   0.4659 \\
            398 &      501 &  -0.73 &  -0.72 &   0.0555 \\
        \hline
        \end{tabular}
    \caption{IceCube effective area from~\cite{ICaeff}}
\end{table}

Here, $\Aeff$ has been averaged over $\Aeff_\mu$ and $\Aeff_{\bar{\mu}}$. Thus, it is not flavor dependent.
Just as with the fluxes, we interpolate this in $\Etrue,\ztrue$ and show the result %TODO aeff interpolation

@@

\begin{table}\label{table:IC_MC}
    \begin{tabular}{lrrrrr}
        \hline \hline
        pdg &      $\Ereco$ [\si{\GeV}] &     $\zreco$ &       $\Etrue$ [\si{\GeV}] &     $\ztrue$ \\
        \hline
         13 &  1665 & -0.645884 &    592 & -0.653421 \\
         13 &   587 & -0.373241 &    342 & -0.424979 \\
        -13 &  1431 & -0.177786 &   1169 & -0.189949 \\
        -13 &   831 & -0.807226 &   1071 & -0.805559 \\
         13 &   988 & -0.370746 &   1861 & -0.367922 \\
         \hline
  \end{tabular}
  \caption{A selection of the data found in~\ref{IC2016}}
\end{table}

@@

The event rate for each bin reads
\begin{align}\label{eq:ICevents}
    N_{ij} &= T \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r \int_0^\pi R(\theta^r,\theta^t) \dd \cos{\theta^t} \int_0^\infty R(E^r,E^t) \dd E^t
    \times \left[ \sum_\beta \phi_\beta^\text{det}  A^\text{eff}_\beta\right]\,,
\end{align}
where $T$ is the live time of the detector. Now this expression handles the Gaussian smearing, but we are not provided systematic error sources, DOM efficiencies, and many more nuisance parameters. To correct this,
we will aim to come as close as possible to the IceCube Monte Carlo, and then normalize with it. That way, we know that our null hypotheses will align while we are free to form additional hypotheses with different 
physics parameters. The binned Monte Carlo events that IceCube used as their null hypothesis in the 2020 sterile analysis is shown in %TODO:MC vs myMC fig.

\begin{figure}[!tb]\label{fig:IC_MC_gpr}
    \begin{center}
       \includegraphics[width=0.4\linewidth]{figures/IC_MC_gpr.pdf}
    \end{center}
    \caption{Relationship between the true and reconstructed muon energy in the IceCube MC sample~\cite{IC2016}}\label{fig:IC_MC_counts}. Shaded area shows the $99.9$th percentile limits predicted by the regressor trained on this set.
 \end{figure}

The latest available data collected and processed by the collaboration contains 305,735 muon track events, collected over eight years~\cite{IC2020}. 
\newpage
\bibliographystyle{nature}
\bibliography{ref.bib}
\end{document}

>>>>

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \input{commands}
% \begin{document}

\section{IceCube}\label{ch:ICmethod}
As the neutrinos have propagated the Earth, they arrive at the South Pole, where they interact with charged lepton in the ice. We now are interested in the effective area $\Aeff$, 
i.e.~the cross-section of the detector that the lepton is exposed to.
$\Aeff$ depends on several parameters, some of them being detector physical volume, $\Etrue$,$\ztrue$ and the neutrino cross-section. Fortunately, the binned $\Aeff$ is provided to us by the collaboration~\cite{ICaeff}.
The data file has the following form

\begin{table}[h]\label{table:aeff}
    \centering
    \begin{tabular}{lrrrrr}
        \hline \hline
        $\Etrue_{min}$ [\si{\GeV}] &     $\Etrue_{max}$ [\si{\GeV}]&   $\ztrue_{min}$ &   $\ztrue_{max}$ &     $\Aeff$ [\si{\metre\squared}] \\
        \hline
             251 &      316 &  -0.92 &  -0.91 &   0.0174 \\
          794300 &  1000000 &  -0.80 &  -0.79 &  69.3600 \\
            3981 &     5012 &  -0.78 &  -0.77 &   3.1490 \\
            1585 &     1995 &  -0.07 &  -0.06 &   0.4659 \\
            398 &      501 &  -0.73 &  -0.72 &   0.0555 \\
        \hline \hline
        \end{tabular}
    \caption{IceCube-86 effective area from~\cite{ICaeff}}
\end{table}

Here, $\Aeff$ has been averaged over $\Aeff_\mu$ and $\Aeff_{\bar{\mu}}$.
Just as with the fluxes, we interpolate this in $\Etrue,\ztrue$ and show the result %TODO aeff interpolation

@@

\begin{table}[h]\label{table:IC_MC}
    \centering
    \begin{tabular}{lrrrrr}
        \hline \hline
        pdg &      $\Ereco$ [\si{\GeV}] &     $\zreco$ &       $\Etrue$ [\si{\GeV}] &     $\ztrue$ \\
        \hline
         13 &  1665 & -0.645884 &    592 & -0.653421 \\
         13 &   587 & -0.373241 &    342 & -0.424979 \\
        -13 &  1431 & -0.177786 &   1169 & -0.189949 \\
        -13 &   831 & -0.807226 &   1071 & -0.805559 \\
         13 &   988 & -0.370746 &   1861 & -0.367922 \\
         \hline \hline
  \end{tabular}
  \caption{A selection of the data found in~\ref{IC2016}}
\end{table}

@@

The event rate for each bin reads
\begin{align}\label{eq:ICevents}
    N_{ij} &= T \sum_\beta\int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r 
    \int_0^\pi R(\theta^r,\theta^t) \dd \cos{\theta^t} \int_0^\infty R(E^r,E^t) \phi_\beta^\text{det}  A^\text{eff}_\beta \dd E^t
    \,,
\end{align}
where $T$ is the live time of the detector. Now this expression handles the Gaussian smearing, but we are not provided systematic error sources, DOM efficiencies, and many more nuisance parameters. To correct this,
we will aim to come as close as possible to the IceCube Monte Carlo, and then normalize with it. That way, we know that our null hypotheses will align while we are free to form additional hypotheses with different 
physics parameters. The binned Monte Carlo events that IceCube used as their null hypothesis in the 2020 sterile analysis is shown in %TODO:MC vs myMC fig.

\begin{figure}[!tb]
    \begin{center}
       \includegraphics[width=0.4\linewidth]{figures/IC_MC_gpr.pdf}
    \end{center}
    \caption{Relationship between the true and reconstructed muon energy in the IceCube MC sample~\cite{IC2016}.
    Shaded area shows the $99.9$th percentile limits predicted by the regressor trained on this set.}\label{fig:IC_MC_gpr}
 \end{figure}

The latest available data collected and processed by the collaboration contains 305,735 muon track events, collected over eight years~\cite{IC2020}. 
%The data has 13 logarithmically spaced bins in $\Ereco \in \[500,9976\] \si{GeV}$, and 20 linear bins in $\zreco \in \[-1,0\]$. The data is shown in Fig.~\ref{fig:IC_data}.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/IC_data.pdf}
    \caption{IceCube track events from~\cite{IC2020}}\label{fig:IC_data}
\end{figure}

\subsection*{Monte Carlo normalization}
Independent researchers outside of the IceCube collaboration will not be able to more persicely
simulate the detector. The IceCube Monte Carlo is a complex and proprietary machinery, so our goal in this 
section is to come as close as we can to their Monte Carlo simulations. After we are confident that 
our code displays the same overall features as the `offical', we normalize our results $N_{ij}^\text{sim}$ as 
\begin{align}\label{eq:MC_norm}
    N_{ij} = \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} N_{ij}^\text{sim}\,.
\end{align}
For each bin $i,j$, we then obtain a correction factor which contains information that we are unable
to obtain or sufficiently incorporate. One example of such information is the systematic errors of the DOMs.
Recent IceCube data releases do not include such information. Since the systematic errors are affecting the 
event count on a bin-by-bin basis, they can in theory drastically modify the binned results. Another example of
an error source what will be remedied by this method is the flux. We are using a fairly simple model of the atmospheric 
flux that excludes prompt flux. The IceCube collaboration use several different flux models which are initialized 
by a parametrization of the cosmic ray flux.\footnote{Included in the cosmic ray models are e.g. the pion to kaon 
ratio, which are often used as a nuisance parameter. By not being able to include this in our error analysis, our method will 
be limited to only consider the overall flux normalization, rather than the components that produce the flux in the first place.}
%TODO: maybe rambling?

In Fig.~\ref{fig:IC_MC_norm}, we present the Icecube Monte Carlo obtained from their 2020 sterile analysis~\cite{IC2020}, along
with our null hypothesis times a constant factor.  We deemed these shapes to be satisfactory, thus allowing us to multiply Eq.~\ref{eq:ICevents} by the 
correction factors of Eq.~\ref{eq:MC_norm}. We now arrive to our final event count
\begin{align}\label{eq:Nth}
    N_{ij} &= \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} T \sum_\beta \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r 
    \int_{E^t_{min}}^{E^t_{max}} R(E^r,E^t) \phi_\beta^\text{det}  A^\text{eff}_\beta\, \dd E^t
    \,,
\end{align}
with $E^t_{min}$, $E^t_{max}$, and $R(E^r,E^t)$ taken from the Gaussian process regressor.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/IC_MC_norm.pdf}
    \caption{IceCube Monte Carlo, binned in $\Ereco$ and $\zreco$. We compare this with our simulations shown as ´Null' in the plots.}\label{fig:IC_MC_norm}
\end{figure} 
%TODO:fill more
% \newpage
% \bibliographystyle{nature}
% \bibliography{ref.bib}
% \end{document}

### 2021-05-28 thesis/tex/SM.tex
update

\documentclass[twocolumn]{article}
\input{commands.tex}
\usepackage{physics,amsmath, amsfonts, siunitx, amssymb}
\usepackage[utf8]{inputenc}
\usepackage{xr-hyper}
%\externaldocument{sterile}

\begin{document}
\section{The Standard Model}\label{ch:SM}
In order to describe the three forces of nature, we gather the mediators of each force -- the vector gauge bosons -- into gauge groups. Each vector boson has one corresponding generator in the group that describes the force. 
The strong charge is mediated by eight gluons, which correspond to the eight generators of $\t{SU}(3)_C$. The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$ and the massless boson $\gamma$, which constitute the generators of $\t{SU}(2)_L$. Finally, the massless gauge boson, the photon, is the generator of $\t{U}(1)_Y$. Together, these groups make up the local symmetry group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. This symmetry group determines the form of the three coupling constants of nature, whose numerical values but be experimentally derived. Thus, the vector bosons are wholly constrained by the symmetry group. However, the scalar bosons and fermions are free as long as they belong to representations of the symmetry group.

The subscript of each group denotes by which mechanism that force is mediated. The gluons mediate the strong force through interactions of color, $C$. The weak force only sees left-handed particles, $L$. And the electroweak interaction that a particle undergoes is determined by its hypercharge, $Y$. For example, the quarks all have a nonzero color and (hyper)charge, so they participate in the strong and electromagnetic interactions. If a given quark is left-handed, it will also feel the weak interaction. The neutrinos on the other hand have neither charge nor color, so they are insivible to both the strong and electromagnetic force.

@@

The leptonic current in Eq.~\ref{eq:j_CC} implies that the lepton number $L_\alpha$ is conserved. 
%\bibliographystyle{apalike}
%\bibliography{ref}
\end{document}

>>>>

% \input{commands}

% \begin{document}
\section{The Standard Model}\label{ch:SM}
In order to describe the three forces of nature, we gather the mediators of each force -- the vector gauge bosons -- into gauge groups. 
Each vector boson has one corresponding generator in the group that describes the force. 
The strong charge is mediated by eight gluons, which correspond to the eight generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$ and the massless boson $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$. Finally, the massless gauge boson, the photon, is the generator of $\text{U}(1)_Y$. 
Together, these groups make up the local symmetry group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This symmetry group determines the form of the three coupling constants of nature, whose numerical values but be experimentally derived. 
Thus, the vector bosons are wholly constrained by the symmetry group. However, the scalar bosons and fermions are free as long as they
 belong to representations of the symmetry group.

The subscript of each group denotes by which mechanism that force is mediated. The gluons mediate the strong force through interactions of color, emphasized with subscript $C$. The weak force only sees left-handed particles, 
which we distinguish wityh the subscript $L$. And the electroweak interaction that a particle undergoes is determined by its hypercharge, $Y$. For example, the quarks all have a nonzero color and (hyper)charge, so they participate in the strong and electromagnetic interactions. If a given quark is left-handed, it will also feel the weak interaction. The neutrinos on the other hand have neither charge nor color, so they are insivible to both the strong and electromagnetic force.

@@

The leptonic current in Eq.~\ref{eq:j_CC} implies that the lepton number $L_\alpha$ is conserved. 
%TODO:more here about SM and lepton mixing?
% \bibliographystyle{apalike}
% \bibliography{ref}
% \end{document}
